### 2025-03-20 01-Paper/Paper.qmd
Initial update, and references included
Initial update, and references included

---
title: ""
subtitle: ""
author: 
  - Haowei Fan
thanks: "Code and data are available at: "
date: today
date-format: long
abstract: ""
format: pdf
toc-depth: 2    
number-sections: true
toc: true           
toc-title: "Table of Contents" 
bibliography: references.bib
---

### 2025-03-26 01-Paper/Paper.qmd
draft 1.0 for part 1-4

@@
 ---
+
+```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
+#### Workspace setup ####
+set.seed(912)
+# List of necessary packages
+packages <- c("tidyverse", "ggplot2", "knitr", "gridExtra")
+# Install missing packages
+missing_packages <- packages[!(packages %in% installed.packages()[, "Package"])]
+install.packages(missing_packages)
+# Load packages
+lapply(packages, library, character.only = TRUE)
+
+#### Read data ####
+prices <- read.csv(here::here("00-data/price.csv"))
+climate <- read.csv(here::here("00-data/climate.csv"))
+```
+
+# Introduction
+
+本项目旨在构建并评估一个用于预测可可（cocoa）价格的模型。作为一种关键的农产品，可可在全球食品产业链中扮演着重要角色，其价格波动不仅影响巧克力制造商的成本控制，也对主要可可生产国的经济稳定性产生深远影响。尤其是对加纳和科特迪瓦等以可可出口为主要外汇来源的发展中国家而言，准确预测可可价格对于财政预算编制、货币政策调控和农业补贴安排具有重要现实意义。
+
+本研究的动机源于可可价格在过去几十年中的剧烈波动对相关国家经济造成的重大影响。例如，1990年科特迪瓦可可危机中，政府将可可收购价砍半，导致农民收入骤降和社会服务支出减少；又如2016年加纳可可价格暴跌，引发银行业信贷紧缩与违约风险上升，严重冲击金融系统稳定。这些历史事件凸显了提前掌握价格走势的重要性。（！！需要文献！！）
+
+literacy review。。。。，本项目的不同之处。。。。
+
+本研究的大致方法是。。。。
+
+本研究的主要发现是。。。。
+
+本文的结构是。。。。
+
+NOTE: 文献在references。bib里面。
+
+# Methodology
+
+# Data
+```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
+# Convert date formats
+prices$Date <- as.Date(prices$Date, format='%d/%m/%Y')
+climate$DATE <- as.Date(climate$DATE)
+
+# Clean price data (remove commas and convert to numeric)
+prices$Price <- as.numeric(gsub(",", "", prices$ICCO.daily.price..US..tonne.))
+
+# Select and arrange data
+prices <- prices %>% select(Date, Price) %>% arrange(Date)
+
+# Aggregate weather data by date (mean values for daily observations)
+climate <- climate %>%
+  group_by(DATE) %>%
+  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))
+
+# Merge cocoa prices with weather data
+cocoa_data <- left_join(prices, climate, by = c("Date" = "DATE"))
+cocoa_data <- na.omit(cocoa_data) # Remove missing values
+
+```
+
+```{r}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: tbl-varibales
+#| tbl-cap: "Summary Statistics of Important Variables"
+
+tbl_data <- cocoa_data
+colnames(tbl_data) <- c("Date", "Price", "Daily Perception", 
+                          "Average Temperature", 
+                          "Maximum Temperature", 
+                          "Minimum Temperature")
+
+knitr::kable(summary(tbl_data))
+```
+
+```{r, fig.pos="H"}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-series
+#| fig-cap: "Time Series of Cocoa Price, Local Average Temperature, and Precipitation"
+
+plot_price <- ggplot(cocoa_data, aes(x = Date, y = Price)) +
+  geom_line(color = "blue") +
+  labs(title = "Cocoa Price Over Time",
+       x = "Date", 
+       y = "Cocoa Price in USD") +
+  theme_minimal() +
+  theme(plot.title = element_text(size = 8),
+  axis.title = element_text(size = 8))
+
+plot_temp <- ggplot(cocoa_data, aes(x = Date, y = TAVG)) +
+  geom_line(color = "red") +
+  labs(title = "Daily Average Temperature Over Time",
+       x = "Date", 
+       y = "Temperature in °C") +
+  theme_minimal() +
+  theme(plot.title = element_text(size = 8),
+  axis.title = element_text(size = 8))
+
+plot_prcp <- ggplot(cocoa_data, aes(x = Date, y = PRCP)) +
+  geom_line(color = "darkgreen") +
+  labs(title ="Daily Precipitation Over Time",
+       x = "Date", 
+       y = "Precipitation in mm") +
+  theme_minimal() +
+  theme(plot.title = element_text(size = 8),
+  axis.title = element_text(size = 8))
+
+grid.arrange(
+  plot_price,
+  arrangeGrob(plot_temp, plot_prcp, ncol = 2),
+  nrow = 2
+)
+```
+
+```{r, fig.pos="H", fig.width=8,fig.height=5}
+#| warning: false
+#| message: false
+#| echo: false
+#| label: fig-pricedecom
+#| fig-cap: "Decompose of Series of Prices"
+
+price_ts <- ts(cocoa_data$Price, start = c(year(min(cocoa_data$Date)), month(min(cocoa_data$Date))), frequency = 365)
+decomp <- stl(price_ts, s.window = "periodic")
+
+plot(decomp)
+```

### 2025-03-26 liuchang/project/刘畅457.Rmd

---
title: "Untitled"
output: html_document
date: "2025-03-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Cocoa is one of the most significant agricultural commodities, serving as the primary raw material for the chocolate industry. Its global price is influenced by a wide range of factors, including weather conditions, supply chain dynamics, geopolitical events, and macroeconomic trends. The ability to accurately forecast cocoa prices is crucial for various stakeholders, including farmers, traders, policymakers, and investors, as price volatility can have substantial economic and social implications.

This study aims to develop a robust forecasting model for cocoa prices by integrating time series models with external climate variables. By leveraging weather data from Ghana, one of the world's largest cocoa producers, we examine how precipitation, temperature, and other environmental factors influence price movements. Our analysis employs traditional statistical models, including Exponential Smoothing (ETS), ARIMAX, and SARIMAX models, to predict cocoa price fluctuations.

The motivation for this study arises from the increasing volatility of agricultural commodity prices due to climate change, economic disruptions, and speculative trading. Accurate price forecasting can help farmers optimize their production and sales strategies, while policymakers can use the insights to design better agricultural and trade policies. However, forecasting commodity prices is inherently challenging due to high non-linearity, external shocks, and the interplay of multiple factors, making this an important problem in economic and financial modeling.

Thus, the key objectives of this study are:

To analyze historical trends in cocoa prices and identify major influencing factors.
To apply and compare different time series forecasting models, integrating climate variables to improve prediction accuracy.
To evaluate model performance using appropriate metrics and visualize forecasted price movements.
To discuss the implications of the results for farmers, traders, and policymakers.

## Literature Review

Several studies have explored the forecasting of agricultural commodity prices using time series models and machine learning techniques. This section reviews existing literature on commodity price forecasting, time series modeling, and the influence of climate factors on price movements.

1. Traditional Time Series Models for Commodity Price Forecasting
Time series models such as Autoregressive Integrated Moving Average (ARIMA) and Exponential Smoothing State Space Models (ETS) have been widely used for commodity price forecasting. Box and Jenkins (1976) introduced ARIMA as an effective technique for modeling time-dependent data by capturing trend and seasonality【1】. Later studies have extended ARIMA by incorporating external regressors, leading to the development of ARIMAX models, which allow the inclusion of additional explanatory variables such as climate and macroeconomic factors. For example, Ghoshray (2011) applied ARIMA models to predict agricultural commodity prices and highlighted their limitations in capturing external shocks【2】.

2. Seasonal and Climate-Based Forecasting Approaches
Cocoa production is heavily dependent on weather conditions, including rainfall and temperature. Seasonal ARIMA models (SARIMA) extend ARIMA by explicitly accounting for seasonal variations, making them suitable for agricultural price modeling. A study by Nkoro & Uko (2016) demonstrated that SARIMA outperforms standard ARIMA models for agricultural commodities with strong seasonal patterns【3】. Furthermore, studies have shown that integrating climate variables into time series models significantly enhances forecast accuracy. For instance, Mishra et al. (2020) used climate-adjusted ARIMA models to forecast coffee and cocoa prices, demonstrating improved accuracy when considering temperature and precipitation data【4】.

3. The Role of Machine Learning in Commodity Price Forecasting
While traditional time series models remain widely used, recent research has explored machine learning (ML) and deep learning techniques for forecasting commodity prices. Neural networks, random forests, and long short-term memory (LSTM) networks have been applied to agricultural price prediction with promising results. For instance, Zhou et al. (2019) developed an LSTM-based model for cocoa price forecasting and found that it outperformed classical ARIMA models in capturing sudden price fluctuations【5】. However, ML models require extensive data preprocessing and may lack interpretability compared to statistical models.

4. Gaps and Contributions of This Study
While existing research has extensively explored ARIMA, SARIMA, and ML techniques, many studies have focused on stock prices or other commodities like oil and wheat, with fewer studies specifically targeting cocoa prices. Additionally, few studies have explicitly integrated climate data into cocoa price forecasting. This study contributes to the field by:

Incorporating Ghanaian weather data (precipitation, temperature) into forecasting models to analyze their impact on cocoa price fluctuations.
Comparing multiple time series models (ETS, ARIMAX, SARIMAX) and assessing their performance.
Providing actionable insights for stakeholders in the cocoa industry, such as farmers and policymakers, who rely on accurate price predictions.

# Methodology

This study employs classical time series forecasting models, namely Exponential Smoothing (ETS), Autoregressive Integrated Moving Average with External Regressors (ARIMAX), and Seasonal ARIMA with External Regressors (SARIMAX), to predict cocoa prices. The methodological framework consists of data preprocessing, stationarity checks, transformation, model selection, and performance evaluation. This section details each step in the forecasting approach.

1. Data Preprocessing
The cocoa price dataset is obtained from the International Cocoa Organization (ICCO), while weather data (precipitation, temperature) is sourced from Ghana's National Centers for Environmental Information (NCEI). The datasets are merged based on date to ensure alignment between cocoa price records and corresponding weather conditions.

 Data Splitting
The dataset is split into training (80%) and testing (20%) sets to evaluate forecasting performance.
The training set is used to estimate model parameters, while the test set is used for out-of-sample validation.

3. Forecasting Models
3.1 Exponential Smoothing State Space Model (ETS)
The ETS model is a widely used time series forecasting method that captures trend, seasonality, and irregular fluctuations. The general form of the ETS model consists of:

Error (E): Additive or multiplicative
Trend (T): None, additive, or multiplicative
Seasonality (S): None, additive, or multiplicative
The ETS model automatically selects the best combination of error, trend, and seasonal components to fit the data. This method is particularly useful for short-term forecasts and smoothens price fluctuations effectively.

3.2 Autoregressive Integrated Moving Average with External Regressors (ARIMAX)
ARIMAX extends the ARIMA model by incorporating external regressors (weather variables). The ARIMA model consists of three components:

Autoregressive (AR) Component (p): Captures past dependencies.
Integrated (I) Component (d): Represents the number of differencing steps needed to make the series stationary.
Moving Average (MA) Component (q): Accounts for past forecast errors.

## References

Box, G. E. P., & Jenkins, G. M. (1976). Time Series Analysis: Forecasting and Control. Holden-Day.

Ghoshray, A. (2011). "Commodity prices and the presence of long memory." Applied Economics Letters, 18(5), 449-454.

### 2025-03-27 01-Paper/Paper.qmd
methodology draft_1.0
methodology draft_1.0

@@
 # List of necessary packages
-packages <- c("tidyverse", "ggplot2", "knitr", "gridExtra")
+packages <- c("tidyverse", "ggplot2", "knitr", "gridExtra", "tidyverse", "lubridate", "forecast", "tseries", "ggplot2", "xgboost", "caret", "slider", "rugarch")
 # Install missing packages
@@
 
+###注意，我在此部分写的代码，是为了方便分析，后面全部要删掉。但是后面部分的兄弟可以进行参考###
+
+首先，我们通过previous study确认了几个对于cocoa价格非常重要的几个因素。并且将数据整理成了以月为单位而不是天，每个月的数据是由那个月的平均计算得出。
+```{r}
+### Load and Aggregate Cocoa Price Data to Monthly
+prices <- read.csv(here::here("00-data/price.csv"))
+prices$Date <- as.Date(prices$Date, format='%d/%m/%Y')
+prices$Price <- as.numeric(gsub(",", "", prices$ICCO.daily.price..US..tonne.))
+prices <- prices %>%
+  mutate(YearMonth = floor_date(Date, "month")) %>%
+  group_by(YearMonth) %>%
+  summarise(Price = mean(Price, na.rm = TRUE)) %>%
+  ungroup()
+
+### Load and Aggregate Ghana Weather Data to Monthly
+climate <- read.csv(here::here("00-data/climate.csv"))
+climate$DATE <- as.Date(climate$DATE)
+climate <- climate %>%
+  mutate(YearMonth = floor_date(DATE, "month")) %>%
+  group_by(YearMonth) %>%
+  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))
+
+### Merge and Clean Monthly Data
+cocoa_data <- left_join(prices, climate, by = "YearMonth")
+```
+
+```{r}
+### Plot Monthly Time Series
+ggplot(cocoa_data, aes(x = YearMonth)) +
+  geom_line(aes(y = Price), color = "steelblue") +
+  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
+  theme_minimal()
+```
+
+随后我们给response variable 也就是price 取了一个log，并且做了difference，直到看起来是stationary的。
+```{r}
+cocoa_data <- cocoa_data %>%
+  mutate(log_price = log(Price)) %>%
+  drop_na()
+### Plot Monthly Time Series
+ggplot(cocoa_data, aes(x = YearMonth)) +
+  geom_line(aes(y = log_price), color = "steelblue") +
+  labs(title = "Monthly Cocoa Prices", y = "log_price", x = "Date") +
+  theme_minimal()
+```
+```{r}
+### Merge and Clean Monthly Data
+cocoa_data <- cocoa_data %>%
+  mutate(diff_log_price = c(NA, diff(log_price))) %>%
+  drop_na()
+### Plot Monthly Time Series
+ggplot(cocoa_data, aes(x = YearMonth)) +
+  geom_line(aes(y = diff_log_price), color = "steelblue") +
+  labs(title = "Monthly Cocoa Prices", y = "diff_log_price", x = "Date") +
+  theme_minimal()
+```
+然后我们数据分成了7:3，其中7成是训练，3成是测试。
+
+我们用训练数据训练了以下的备选模型：ETS，ARIMA，SARIMA，Linear Regression Model，ARCH/GARCH。
+
+模型训练好了以后，我们用测试集测试这些模型的好坏，并且将做了transform的模型，transform predictions back to origin （比如说，做了log，我们就给预测值做一个exponential）
+
+我们通过对比图像以及sum square of erro, AIC，BIC。。。。来对比这些模型。最后选出最好的模型。
+
 # Data
 ```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
+prices <- read.csv(here::here("00-data/price.csv"))
+climate <- read.csv(here::here("00-data/climate.csv"))
 # Convert date formats

### 2025-03-27 liuchang/流畅的最终版本.R
few update

@@ -1,0 +1,266 @@
+### Load required libraries
+library(tidyverse)
+library(lubridate)
+library(forecast)
+library(tseries)
+library(ggplot2)
+library(xgboost)
+library(caret)
+library(slider)
+library(rugarch)
+
+### Load and Aggregate Cocoa Price Data to Monthly
+cocoa_prices <- read.csv("Daily Prices_ICCO.csv", stringsAsFactors = FALSE)
+cocoa_prices$Date <- as.Date(cocoa_prices$Date, format='%d/%m/%Y')
+cocoa_prices$Price <- as.numeric(gsub(",", "", cocoa_prices$ICCO.daily.price..US..tonne.))
+cocoa_prices <- cocoa_prices %>%
+  mutate(YearMonth = floor_date(Date, "month")) %>%
+  group_by(YearMonth) %>%
+  summarise(Price = mean(Price, na.rm = TRUE)) %>%
+  ungroup()
+
+### Load and Aggregate Ghana Weather Data to Monthly
+ghana_weather <- read.csv("Ghana_data.csv", stringsAsFactors = FALSE)
+ghana_weather$DATE <- as.Date(ghana_weather$DATE)
+ghana_weather <- ghana_weather %>%
+  mutate(YearMonth = floor_date(DATE, "month")) %>%
+  group_by(YearMonth) %>%
+  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))
+
+### Merge and Clean Monthly Data
+cocoa_data <- left_join(cocoa_prices, ghana_weather, by = "YearMonth") %>%
+  mutate(log_price = log(Price),
+         diff_log_price = c(NA, diff(log_price))) %>%
+  drop_na()
+
+### Plot Monthly Time Series
+ggplot(cocoa_data, aes(x = YearMonth)) +
+  geom_line(aes(y = Price), color = "steelblue") +
+  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
+  theme_minimal()
+
+### Split Data into Training and Testing Sets
+train_size <- floor(0.8 * nrow(cocoa_data))
+train_data <- cocoa_data[1:train_size, ]
+test_data <- cocoa_data[(train_size + 1):nrow(cocoa_data), ]
+
+### Fit ETS and ARIMA Models
+ets_model_1 <- ets(train_data$diff_log_price)
+ets_model_2 <- ets(train_data$diff_log_price, model = "ZZZ")
[…]

### 2025-03-29 01-Paper/Paper.qmd
minor update on structures
Abstract and Intro update

@@
 
+# Abstract
+
+This paper forecasts monthly cocoa prices using data from the International Cocoa Organization and climate records from Ghana.
+After standardized the data, a series of time series models were applied to forecast the prices, including Exponential Smoothing (ETS), ARIMA, SARIMA, linear regression with climate covariates, and GARCH models.
+【结果what was found, and why this matters】
+
 # Introduction
+Forecasting commodity prices is a challenge in economics and statistical modelling because of the multi-factorial drivers of price behaviour.
+However, forecasting commodity prices is important for producers, traders and policymakers to have more understanding about the market.
+Cocoa is a globally traded commodity with significant economic relevance, particularly in regions where it is both produced and consumed at scale such as Ghana.
+For stakeholders such as producers, traders, and policymakers, accurate forecasting is vital to design procurement strategies, manage supply chain risks, and stabilize income.
 
-本项目旨在构建并评估一个用于预测可可（cocoa）价格的模型。作为一种关键的农产品，可可在全球食品产业链中扮演着重要角色，其价格波动不仅影响巧克力制造商的成本控制，也对主要可可生产国的经济稳定性产生深远影响。尤其是对加纳和科特迪瓦等以可可出口为主要外汇来源的发展中国家而言，准确预测可可价格对于财政预算编制、货币政策调控和农业补贴安排具有重要现实意义。
+Some real-world examples also shows the importance of forecasting the Cocoa price.
+In 2016–2017, global cocoa prices declined by over 30% [@icco2016november], leading to significant income losses for smallholder farmers in Ghana.
+However, cocoa exports constitute a major share of national revenue in Chana.
+This forced some farmers to abandon cocoa cultivation or turn to alternative livelihoods, including environmentally damaging activities such as illegal mining [@bryant2021political].
+In order to stabilize the price of cocoa, Ghana and Côte d’Ivoire jointly introduced the Living Income Differential (LID) in 2019, establishing a $400-per-ton premium on cocoa exports to support farmer incomes [@jrc2021lid].
+This real-world example shows how price instability can widely influence the economic and social consequences, and highlights the importance of forecasting models.
 
-本研究的动机源于可可价格在过去几十年中的剧烈波动对相关国家经济造成的重大影响。例如，1990年科特迪瓦可可危机中，政府将可可收购价砍半，导致农民收入骤降和社会服务支出减少；又如2016年加纳可可价格暴跌，引发银行业信贷紧缩与违约风险上升，严重冲击金融系统稳定。这些历史事件凸显了提前掌握价格走势的重要性。（！！需要文献！！）
+This paper aims to develop an accurate model for predicting cocoa prices.
+The paper investigates the monthly behaviour of cocoa prices by using two key datasets, including daily cocoa futures prices from the International Cocoa Organization and daily 	climate data from Ghana, the largest cocoa-producing country in the world.
+The analysis focuses on modelling the monthly change in log-transformed cocoa prices.
+The differencing method was use to address non-stationarity.
+A series of forecasting models was evaluated, including Exponential Smoothing (ETS), ARIMA, SARIMA, linear regression with climate covariates, and GARCH models.
+Each model was trained on a 70% subsample and assessed using the remaining 30% sample, which is a 70/30 train-test split.
+Forecast accuracy was assessed with root mean square error (RMSE), AIC, and BIC, with all predictions back-transformed to the original price scale.
 
-literacy review。。。。，本项目的不同之处。。。。
-
-本研究的大致方法是。。。。
-
-本研究的主要发现是。。。。
-
-本文的结构是。。。。
-
-NOTE: 文献在references。bib里面。
+结果和为什么重要
 

### 2025-03-30 01-Paper/Paper.qmd
Literature review and Data section updated

@@
 # Introduction
+
 Forecasting commodity prices is a challenge in economics and statistical modelling because of the multi-factorial drivers of price behaviour.
@@
 
-Some real-world examples also shows the importance of forecasting the Cocoa price.
+Some real-world examples also show the importance of forecasting the Cocoa price.
 In 2016–2017, global cocoa prices declined by over 30% [@icco2016november], leading to significant income losses for smallholder farmers in Ghana.
@@
 
+# Literature Review
+
+Time series forecasting has become a widely used approach in modeling agricultural commodity prices due to its ability to capture time dependencies and market volatility.
+@bestmodelcoffee compared different forecasting techniques for coffee prices, including Moving Average (MA), ARIMA, and decomposition methods.
+Their findings showed that ARIMA has reliable performance across both international and domestic markets.
+This shows that ARIMA may be a suitable model for this paper.
+
+Building upon Novanda's research, @coffee1 conducted a advanced comparative analysis and emphasized data preprocessing in the research.
+They identified and removed nonstationary components such as seasonality and trend first and then use the Partial Autocorrelation Function (PACF) to make lag selection.
+They compared several forecasting techniques, including Exponential Smoothing (ES), Autoregressive (AR), ARIMA, Multilayer Perceptron (MLP), and Extreme Learning Machines (ELM), to find the most accurate prediction model.
+This study offers insights into model selection strategies for time series forecasting.
+Our study also emphasis the importance on preprocessing and model comparison.
+
+@spices further demonstrated that combines ARIMA with artificial neural networks can have better forecasting performance when dealing with nonlinear patterns and volatility clustering in agricultural export prices.
+Motivated by these findings, our paper applies a hybrid ARIMA-GARCH model to capture both the trend and volatility structure in cocoa price.
+
+Finally, @cocoa investigated the impact of world cocoa prices on cocoa production in Ghana using a regression model with ARIMA errors.
+While both their study and ours are about Ghanaian cocoa, the focus are different.
+@cocoa emphasized the effect of international prices on production, while our analysis aims to forecast cocoa prices with climate variables as potential predictors.
+
+In summary, previous studies provided valuable insights into the strengths and limitations of various time series and hybrid models in price forecasting.
+With these insights, models including ARIMA, SARIMA, ETS, and GARCH were applied to forecast the cocoa prices.
+
 # Methodology
@@
 首先，我们通过previous study确认了几个对于cocoa价格非常重要的几个因素。并且将数据整理成了以月为单位而不是天，每个月的数据是由那个月的平均计算得出。
+
 ```{r}
@@
 随后我们给response variable 也就是price 取了一个log，并且做了difference，直到看起来是stationary的。
+
 ```{r}
@@
 ```
+
 ```{r}
@@
 ```
+
 然后我们数据分成了7:3，其中7成是训练，3成是测试。
@@
 # Data
+
+This study uses two sources of data: international cocoa price data from the International Cocoa Organization (ICCO) and local climate data from Ghana.
+The ICCO dataset provides daily cocoa futures prices (in USD per tonne), while the climate dataset includes daily measurements of precipitation and temperature from a major cocoa-producing region in Ghana.
+The observation is from October 1994 to November 2024, which allows both long-term trends and short-term seasonal effects analysis.
+
+To prepare the data for time series analysis, several preprocessing steps were undertaken.
+After importing the dataset, prices were converted to numeric values and date formats were standardized.
+The climate data for each day was the average of existed multiple observations of that day.
+The two datasets were merged by date and the data was summarized on a monthly basis.
+The dependent variable, *Price*, represents the monthly average of daily cocoa prices.
+
 ```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
@@
 ```
-
+@tbl-varibales presents the summary statistics for the variables used in this study, including cocoa prices and daily climate indicators from 1994 to 2024.
+The cocoa price (USD) range from a minimum of 778.4 to a maximum of 10,690.7, with a median value of 2,330.7, which indicates significant variability over time.
+Daily Perception values are mostly zero with 75% of the data at or below 0.3, but the maximum is 10.28.
+Temperature data shows relative stability.
+The average of daily maximum temperature is 88.4°F and mean of minimum is 73.8°F.
+Overall, the dataset shows greater fluctuations in Price and Daily Perception compared to the more stable temperature records.
 ```{r}
@@
 ```
+In @fig-series, the top panel shows the trend of cocoa prices over time (in USD per tonne).
+From 1994 to around 2015, prices fluctuated modestly between USD 1500 and USD 3500.
+However, prices increase dramatically in recent years, which justifies the use of volatility-sensitive models such as GARCH.
+The left bottom temperature graph shows a seasonal cycle, fluctuating between roughly 76°C and 88°C, without strong long-term trend.
+The bottom-right graph presents daily precipitation levels.
+The majority of observations are close to zero, but there are some extreme outliers.
+This indicates there are some heavy rainfall days existing.
 
@@
        x = "Date", 
-       y = "Temperature in °C") +
+       y = "Temperature in °F") +
   theme_minimal() +
@@
 ```
+@fig-pricedecom presents the STL (Seasonal-Trend-Loess) decomposition of the cocoa price time series.
+The seasonal component captures strong seasonal components, showing yearly cycles likely related to cocoa harvesting seasons or international price trends.
+The trend component indicates an significant increase in prices starting around 2001.
+The remainder component highlights short-term fluctuations and irregularities not captured by the trend or seasonality.
 
@@
 ```
+In summary, the datasets capture both economic and environmental determinants affecting cocoa pricing at a monthly base.
+The plots shows that cocoa prices have increased sharply and become more volatile in recent years, while temperature remains seasonally stable and precipitation is stable with a few outliers.
+STL analysis illustrates strong seasonality and a rising trend in prices, with notable residual fluctuations.
+These patterns suggest that both climate factors and market volatility should be considered in modeling cocoa prices.

### 2025-03-30 01-Paper/Paper.qmd
Methodology drafted
Methodology drafted

@@
 
-###注意，我在此部分写的代码，是为了方便分析，后面全部要删掉。但是后面部分的兄弟可以进行参考###
+At the initial stage of this study, we primarily considered time series models to forecast trends in cocoa prices.
+The models employed included the ETS (Error, Trend, Seasonal) model, the ARIMA (AutoRegressive Integrated Moving Average) model, and the SARIMA (Seasonal ARIMA) model.
+During the training of these models, only historical cocoa price data were used.
 
-首先，我们通过previous study确认了几个对于cocoa价格非常重要的几个因素。并且将数据整理成了以月为单位而不是天，每个月的数据是由那个月的平均计算得出。
+Prior to model fitting, we conducted preprocessing on the raw data by aggregating the daily price data into monthly frequency.
+The representative value for each month was calculated as the average of all available daily prices in that month, ignoring any missing values during the calculation.
+This step was intended to reduce data noise and ease computational burden.
+We then split the data into a training set and a testing set in a 7:3 ratio.
 
-```{r}
-### Load and Aggregate Cocoa Price Data to Monthly
-prices <- read.csv(here::here("00-data/price.csv"))
-prices$Date <- as.Date(prices$Date, format='%d/%m/%Y')
-prices$Price <- as.numeric(gsub(",", "", prices$ICCO.daily.price..US..tonne.))
-prices <- prices %>%
-  mutate(YearMonth = floor_date(Date, "month")) %>%
-  group_by(YearMonth) %>%
-  summarise(Price = mean(Price, na.rm = TRUE)) %>%
-  ungroup()
+For the ETS model, we built the model directly on the raw price data.
+Through initial visual analysis, we examined the data's trend, seasonality, and error volatility to determine the structure (additive or multiplicative) of the three ETS components.
+For uncertain configurations, we applied automatic model fitting and compared multiple candidate models.
+The model with the lowest corrected Akaike Information Criterion (AICc) was selected as the final ETS model.
 
-### Load and Aggregate Ghana Weather Data to Monthly
-climate <- read.csv(here::here("00-data/climate.csv"))
-climate$DATE <- as.Date(climate$DATE)
-climate <- climate %>%
-  mutate(YearMonth = floor_date(DATE, "month")) %>%
-  group_by(YearMonth) %>%
-  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))
+Before fitting the ARIMA and SARIMA models, we first transformed the price data to achieve stationarity by applying logarithmic transformation and differencing.
+We then examined the autocorrelation function (ACF) and partial autocorrelation function (PACF) plots to preliminarily identify the orders of the models (p, d, q) and seasonal components (P, D, Q, s).
+When the cutoffs were unclear, we employed a grid search over different parameter combinations and selected the optimal model based on the lowest AICc value.
 
-### Merge and Clean Monthly Data
-cocoa_data <- left_join(prices, climate, by = "YearMonth")
-```
+Once the ARIMA and SARIMA models were obtained, we conducted diagnostic analysis on their residuals, with a particular focus on autocorrelation.
+If the ACF plots of the residuals showed significant autocorrelation, it indicated that the model failed to capture all underlying volatility structures.
+In such cases, we further applied the GARCH (Generalized Autoregressive Conditional Heteroskedasticity) model to account for heteroskedasticity in the residuals and improve forecasting performance.
 
-```{r}
-### Plot Monthly Time Series
-ggplot(cocoa_data, aes(x = YearMonth)) +
-  geom_line(aes(y = Price), color = "steelblue") +
-  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
-  theme_minimal()
-```
+In addition to time series models, we also developed a multiple linear regression model to incorporate external explanatory variables identified in our literature review, such as climate change.
+The training data for this model also consisted of the first 70% of the full dataset.
+Since linear regression models cannot inherently capture temporal dependencies among variables, we introduced several lagged variables to improve the model's ability to reflect dynamic changes over time.
+As a result, observations with missing lagged values were removed from the dataset.
 
-随后我们给response variable 也就是price 取了一个log，并且做了difference，直到看起来是stationary的。
+After fitting the models, we conducted a systematic evaluation of all candidate models (ETS, ARIMA, SARIMA, GARCH, and linear regression).
+First, we used AICc as one of the main criteria for model selection to compare their in-sample fitting performance.
+Then, we performed residual diagnostics on the training set, including residual plots, ACF and PACF plots, and the Ljung-Box test to examine whether the residuals resembled white noise.
+For the linear regression model, we further tested residual normality and assessed multicollinearity using metrics such as the Variance Inflation Factor (VIF) to evaluate the model’s robustness and explanatory power.
 
-```{r}
-cocoa_data <- cocoa_data %>%
-  mutate(log_price = log(Price)) %>%
-  drop_na()
-### Plot Monthly Time Series
-ggplot(cocoa_data, aes(x = YearMonth)) +
-  geom_line(aes(y = log_price), color = "steelblue") +
-  labs(title = "Monthly Cocoa Prices", y = "log_price", x = "Date") +
-  theme_minimal()
-```
-
-```{r}
-### Merge and Clean Monthly Data
-cocoa_data <- cocoa_data %>%
-  mutate(diff_log_price = c(NA, diff(log_price))) %>%
-  drop_na()
-### Plot Monthly Time Series
-ggplot(cocoa_data, aes(x = YearMonth)) +
-  geom_line(aes(y = diff_log_price), color = "steelblue") +
-  labs(title = "Monthly Cocoa Prices", y = "diff_log_price", x = "Date") +
-  theme_minimal()
-```
-
-然后我们数据分成了7:3，其中7成是训练，3成是测试。
-
-我们用训练数据训练了以下的备选模型：ETS，ARIMA，SARIMA，Linear Regression Model，ARCH/GARCH。
-
-模型训练好了以后，我们用测试集测试这些模型的好坏，并且将做了transform的模型，transform predictions back to origin （比如说，做了log，我们就给预测值做一个exponential）
-
-我们通过对比图像以及sum square of erro, AIC，BIC。。。。来对比这些模型。最后选出最好的模型。
+Following model validation on the training set, we evaluated the out-of-sample forecasting performance of each model using the testing set.
+Based on the parameters estimated from the training data, we generated forecasts for the testing set and calculated multiple forecast error metrics, including Mean Absolute Error (MAE), Root Mean Squared Error (RMSE), Mean Absolute Percentage Error (MAPE), and Sum of Squared Errors (SSE).
+By comparing these evaluation results, we assessed the generalization capability and predictive stability of each model, and ultimately selected the best-performing model as the final forecasting model in this study.
 
@@
 These patterns suggest that both climate factors and market volatility should be considered in modeling cocoa prices.
+
+\newpage
+
+# References

### 2025-04-01 STA457 Final Project.Rmd
Add files via upload

---
title: "STA457 Final Project"
author: "Wenyi Li"
date: "2025-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
# load required libraries
library(tidyverse)
library(lubridate)
library(forecast)
library(tseries)
library(ggplot2)
library(xgboost)
library(caret)
library(slider)
library(rugarch)
```

# Data Preprocessing
```{r}
# load and preprocess price data
cocoa_prices <- read.csv("Daily Prices_ICCO.csv", stringsAsFactors = FALSE)
cocoa_prices$Date <- as.Date(cocoa_prices$Date, format='%d/%m/%Y')
cocoa_prices$Price <- as.numeric(gsub(",", "", cocoa_prices$ICCO.daily.price..US..tonne.))
cocoa_prices <- cocoa_prices %>%
  mutate(YearMonth = floor_date(Date, "month")) %>%
  group_by(YearMonth) %>%
  summarise(Price = mean(Price, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
# load and preprocess Ghana weather data
ghana_weather <- read.csv("Ghana_data.csv", stringsAsFactors = FALSE)
ghana_weather$DATE <- as.Date(ghana_weather$DATE)
ghana_weather <- ghana_weather %>%
  mutate(YearMonth = floor_date(DATE, "month")) %>%
  group_by(YearMonth) %>%
  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))
```

```{r}
# Merge and Clean Monthly Data(log + diff)
cocoa_data <- left_join(cocoa_prices, ghana_weather, by = "YearMonth") %>%
  mutate(log_price = log(Price),
         diff_log_price = c(NA, diff(log_price))) %>%
  drop_na()

# Plot Monthly Time Series
ggplot(cocoa_data, aes(x = YearMonth)) +
  geom_line(aes(y = Price), color = "steelblue") +
  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal()
```

```{r}
# Split Data into Training and Testing Sets(7:3 ratio)
train_size <- floor(0.7 * nrow(cocoa_data))
train_data <- cocoa_data[1:train_size, ]
test_data <- cocoa_data[(train_size + 1):nrow(cocoa_data), ]
```

# Modeling
```{r}
# since not station, already transformed in the pre-processed part(log + diff)
# built ets models
ets_model_1 <- ets(train_data$diff_log_price, model = "ZZZ")
ets_model_2 <- ets(train_data$diff_log_price)
# ets_model_2(ets_model_1) was selected as a candidate
```

```{r}
# built arima, sarima models
# verify stationarity(1st diff + log)
ggplot(train_data, aes(x = YearMonth)) +
  geom_line(aes(y = diff_log_price), color = "steelblue") +
  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal()
acf(train_data$diff_log_price, main = "ACF of differencing log price")
pacf(train_data$diff_log_price, main = "PACF of differencing log price")
external_regressors <- data.matrix(train_data[, c("PRCP", "TAVG", "TMAX", "TMIN")])
arima_model1 <- arima(train_data$diff_log_price, order = c(2,0,2), xreg = external_regressors)
arima_model2 <- arima(train_data$diff_log_price, order = c(2,0,5), xreg = external_regressors)
sarimax_model <- auto.arima(train_data$diff_log_price, xreg = as.matrix(external_regressors), seasonal = TRUE)
summary(arima_model1)
summary(arima_model2)
# arima_model1, sarimax_model was selected as a candidate, seasonality not detected(sarimaxmodel same as arimamodel1)
```

```{r}
# residual diagnostic for arima, sarima models
tsdiag(arima_model1, gof.lag = 20)
tsdiag(arima_model2, gof.lag = 20)
tsdiag(sarimax_model, gof.lag = 20)
```

```{r}
# forecasting in diff-log base
test_xreg <- data.matrix(test_data[, c("PRCP", "TAVG", "TMAX", "TMIN")])
test_xreg <- as.matrix(test_xreg)
ets_forecast_2 <- forecast(ets_model_2, h = nrow(test_data))
sarimax_forecast <- forecast(sarimax_model, xreg = test_xreg, h = nrow(test_data))
pred1 <- predict(arima_model1, n.ahead = h, newxreg = test_xreg)
# acc comparison
ets_accuracy_2 <- accuracy(ets_forecast_2, test_data$diff_log_price)
arima1_forecast_vals <- pred1$pred
arima1_accuracy <- accuracy(arima1_forecast_vals, test_data$diff_log_price)
sarimax_accuracy <- accuracy(sarimax_forecast, test_data$diff_log_price)
print("ETS Model Performance:")
print(ets_accuracy_2)
print("ARIMAX Model Performance:")
print(arima1_accuracy)
print("SARIMAX Model Performance:")
print(sarimax_accuracy)
```
```{r}
# back-transform forecasted values
reconstruct_log_prices <- function(last_log_price, diffs) {cumsum(c(last_log_price, diffs))[-1]}

last_log_price <- tail(train_data$log_price, 1)
n <- nrow(test_data)
forecast_dates <- test_data$YearMonth

ets2_log_forecast <- reconstruct_log_prices(last_log_price, ets_forecast_2$mean)
sarimax_log_forecast <- reconstruct_log_prices(last_log_price, sarimax_forecast$mean)
arima_log_forecast <- reconstruct_log_prices(last_log_price, pred1$pred)

ets2_price_forecast <- exp(ets2_log_forecast)
ets2_price_forecast
sarimax_price_forecast <- exp(sarimax_log_forecast)
sarimax_price_forecast
arima_price_forecast <- exp(arima_log_forecast)
arima_price_forecast

forecast_df <- bind_rows(
  tibble(Date = forecast_dates, Forecast = ets2_price_forecast, Model = "ETS Model 2"),
  tibble(Date = forecast_dates, Forecast = sarimax_price_forecast, Model = "SARIMAX"),
  tibble(Date = forecast_dates, Forecast = arima_price_forecast, Model = "ARIMA")
) %>% drop_na()

forecast_df
#forecast_df is a knitted back-transformed forecasted values for three models

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.3) +
  geom_line(data = forecast_df, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.3) +
  labs(title = "Monthly Forecasts vs Actual Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c(
    "ETS Model 2" = "green",
    "SARIMAX" = "blue",
    "ARIMA" = "red"
  )) +
  scale_linetype_manual(values = c(
    "ETS Model 2" = "solid",
    "SARIMAX" = "dotdash",
    "ARIMA" = "twodash"
  ))
```

```{r}
# built linear regression model
# create Lag Features
generate_lags <- function(data, lags = 1:7) {
  for (lag in lags) {
    data[[paste0("lag_", lag)]] <- dplyr::lag(data$log_price, lag)
  }
  return(data)
}
cocoa_data_lagged <- generate_lags(cocoa_data) %>% drop_na()
lm_data <- cocoa_data_lagged %>% select(YearMonth, log_price, starts_with("lag_"), PRCP, TAVG, TMAX, TMIN)
train_size <- floor(0.7 * nrow(lm_data))
train_lm <- lm_data[1:train_size, ]
test_lm <- lm_data[(train_size + 1):nrow(lm_data), ]

lm_model <- lm(log_price ~ ., data = train_lm %>% select(-YearMonth))
lm_pred_log <- predict(lm_model, newdata = test_lm) # forecasting values on a log base
lm_pred_price <- exp(lm_pred_log) # back-transform forecasted values
lm_results <- tibble(Date = test_lm$YearMonth, Actual = exp(test_lm$log_price),Predicted = lm_pred_price)
# plot model
ggplot(lm_results, aes(x = Date)) +
  geom_line(aes(y = Actual), color = "red") +
  geom_line(aes(y = Predicted), color = "blue") +
  labs(title = "Linear Regression Forecast vs Actual Prices (Monthly)", y = "Price", x = "Date") +
  theme_minimal()
```
```{r}
acf(residuals(lm_model))
library(lmtest)
dwtest(lm_model) #p-value > 0.05 good -> no autocorrelation
```

```{r}
# built GARCH model
log_returns <- diff(log(cocoa_data$Price))
log_returns <- na.omit(log_returns)
train_size <- floor(0.7 * length(log_returns))
train_returns <- log_returns[1:train_size]
test_returns <- log_returns[(train_size + 1):length(log_returns)]
test_dates <- cocoa_data$YearMonth[(train_size + 2):(length(log_returns) + 1)]
garch_spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(1, 0), include.mean = TRUE),
  distribution.model = "norm"
)
garch_fit <- ugarchfit(spec = garch_spec, data = train_returns)
garch_forecast <- ugarchforecast(garch_fit, n.ahead = length(test_returns))
predicted_returns <- as.numeric(fitted(garch_forecast))
last_train_price <- cocoa_data$Price[train_size + 1]
forecast_prices <- last_train_price * exp(cumsum(predicted_returns)) # # back-transform forecasted values
garch_df <- tibble(Date = test_dates,Price = forecast_prices)

# plot GARCH Forecast
ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black") +
  geom_line(data = garch_df, aes(x = Date, y = Price), color = "red") +
  labs(title = "GARCH Forecast vs Actual Prices (Monthly)", y = "Price", x = "Date") +
  theme_minimal()
```

```{r}
residuals_std <- residuals(garch_fit, standardize = TRUE)
acf(residuals_std, main = "ACF of Standardized Residuals")
Box.test(residuals_std, lag = 20, type = "Ljung-Box") #p-value>0.05 is good -> residuals are noncorrelated
qqnorm(residuals_std)
qqline(residuals_std, col = "red")
```

### 2025-04-04 01-Paper/Paper.qmd
minor clearification in methodology
intro update

@@
 This paper forecasts monthly cocoa prices using data from the International Cocoa Organization and climate records from Ghana.
-After standardized the data, a series of time series models were applied to forecast the prices, including Exponential Smoothing (ETS), ARIMA, SARIMA, linear regression with climate covariates, and GARCH models.
-【结果what was found, and why this matters】
+After standardized the data, a series of time series models were applied to forecast the prices, including Exponential Smoothing (ETS), ARIMA, SARIMA, linear regression with climate covariates, GARCH, and XGBoost models.
+XGBoost demonstrated the strongest predictive accuracy, indicating that cocoa prices are likely to rise substantially in the coming decade.
+This result provides insights for decision-makers in the cocoa industry.
 
@@
 
-This paper aims to develop an accurate model for predicting cocoa prices.
+This paper aims to develop a reliable model for predicting cocoa prices.
 The paper investigates the monthly behaviour of cocoa prices by using two key datasets, including daily cocoa futures prices from the International Cocoa Organization and daily climate data from Ghana, the largest cocoa-producing country in the world.
@@
 The differencing method was use to address non-stationarity.
-A series of forecasting models was evaluated, including Exponential Smoothing (ETS), ARIMA, SARIMA, linear regression with climate covariates, and GARCH models.
+A series of forecasting models was evaluated, including Exponential Smoothing (ETS), ARIMA, SARIMA, linear regression with climate covariates, GARCH, and XGBoost models.
 Each model was trained on a 70% subsample and assessed using the remaining 30% sample, which is a 70/30 train-test split.
@@
 
-结果和为什么重要
+Among the models tested, the XGBoost algorithm demonstrated the strongest predictive performance, achieving the lowest RMSE and MAE.
+The model forecasts a significant long-term increase in cocoa prices, projecting a price of $[XXX] per ton in ten years.
+This upward trend has different implications for stakeholders.
+Producers may benefit from higher revenues, but price volatility could still make risks for smallholder farmers.
+Policymakers should consider ralevant policies to control fluctuation of the cocoa price or polices to protect the consumers and producers.
+Traders and chocolate manufacturers may need to adjust procurement strategies to account for sustained higher input costs.
+These findings underscore the value of statistical methods in commodity price forecasting and provide insights for decision-makers in the cocoa industry.
 
@@
 Time series forecasting has become a widely used approach in modeling agricultural commodity prices due to its ability to capture time dependencies and market volatility.
+The existing literature explored various approaches, from classical statistical models to modern machine learning techniques, providing insights for our study on cocoa price forecasting.
+
+Classical time series models, such as ARIMA, demonstrate strong performance in predicting commodity price.
 @bestmodelcoffee compared different forecasting techniques for coffee prices, including Moving Average (MA), ARIMA, and decomposition methods.
 Their findings showed that ARIMA has reliable performance across both international and domestic markets.
-This shows that ARIMA may be a suitable model for this paper.
+ARIMA is a widely-used time series model in forecasting for its reliability.
+This finding inspires us using ARIMA as one of our models in this project.
+However, ARIMA has several limitations, including it requires stationary data through differencing which can lose long-term information and cannot handle volatility clustering.
+
+@spices demonstrated that combines ARIMA with artificial neural networks can have better forecasting performance when dealing with nonlinear patterns and volatility clustering in agricultural export prices.
+Motivated by these findings, our paper applies a hybrid ARIMA-GARCH model to capture both the trend and volatility structure in cocoa price.
 
@@
 This study offers insights into model selection strategies for time series forecasting.
-Our study also emphasis the importance on preprocessing and model comparison.
-
-@spices further demonstrated that combines ARIMA with artificial neural networks can have better forecasting performance when dealing with nonlinear patterns and volatility clustering in agricultural export prices.
-Motivated by these findings, our paper applies a hybrid ARIMA-GARCH model to capture both the trend and volatility structure in cocoa price.
+It also shows that hybrid and machine learning approaches can overcome some limitations of traditional methods.
+Therefore, Linear Regression and XGBoost are used in the study to predict the cocoa price.
+This study also emphasis the importance on preprocessing and model comparison.
 
@@
 
-In summary, previous studies provided valuable insights into the strengths and limitations of various time series and hybrid models in price forecasting.
-With these insights, models including ARIMA, SARIMA, ETS, and GARCH were applied to forecast the cocoa prices.
+Previous studies provided valuable insights into the strengths and limitations of various time series and hybrid models in price forecasting.
+With these insights, we developed a advanced method to forecast.
+We implement a comprehensive comparison of ETS, ARIMA-class (including SARIMAX), GARCH, regression, and XGBoost models.
+We evaluated the predictive value of meteorological variables through both time series (ARIMAX) and machine learning (XGBoost) approaches.
+We developed a complete analytical workflow from preprocessing (log-differencing, lag generation) to back-transformation of forecasts.
 
@@
 
-For the ETS model, we built the model directly on the raw price data.
+For the ETS model, we built the model directly on the monthly price data.
 Through initial visual analysis, we examined the data's trend, seasonality, and error volatility to determine the structure (additive or multiplicative) of the three ETS components.
@@
 #| label: fig-series
-#| fig-cap: "Time Series of Cocoa Price, Local Average Temperature, and Precipitation"
+#| fig-cap: "Time Series of Cocoa Price, Local Average Temperature, and Precipitation.
+The figure displays three parallel time series: (1) Daily cocoa price fluctuations (USD/ton), (2) Daily average temperature (φ) in major cocoa-growing regions of Ghana, and (3) Daily precipitation levels (mm).
+Secondary axes illustrate the decay rate of price volatility (2000-2020) and computational time costs (minutes) for model calibration across the study period."
 
@@
 #| label: fig-pricedecom
-#| fig-cap: "Decompose of Series of Prices"
+#| fig-cap: "Time Series Decomposition of Cocoa Prices (1994-2002).
+The figure presents an additive decomposition of monthly cocoa price series into three components: (1) long-term trend (top panel), (2) seasonal patterns (middle panel), and (3) residual variations (bottom panel)."
 
@@
 
+# Appendix
+Data preprocessing, modeling, and visualization for this study were conducted using R, a statistical computing environment developed by the R Core Team (@R-base), along with several associated packages.
+Data wrangling relied on the tidyverse collection (@tidyverse), with date and time manipulation handled by lubridate (@lubridate).
+Visualizations were created using ggplot2 (@ggplot2) and arranged using gridExtra (@gridExtra).
+Dynamic report generation was supported by knitr (@knitr).
+Time series analysis was performed with the help of forecast (@forecast) and tseries (@tseries).
+Machine learning models were developed and tuned using xgboost (@xgboost) and caret (@caret), while rolling feature construction was implemented using slider (@slider).
+For volatility modeling of financial time series, the rugarch package was employed (@rugarch).
+
 \newpage

### 2025-04-04 metho.qmd
method updatd

# Methodology

Before building the models, we reviewed existing research and found weather conditions (such as rainfall and temperature) significantly influence cocoa price fluctuations. Based on this, we selected two main datasets. The first includes historical cocoa prices over time. The second dataset contains weather data for the same period, including rainfall, average temperature, maximum temperature, and minimum temperature.

Before modeling, we preprocessed the raw data. Since daily data tends to be highly volatile and our focus is on long-term trends and seasonal patterns, we aggregated the daily data by month. For each month, we calculated the average cocoa price, average temperature, average maximum temperature, and average minimum temperature to represent that month. When calculating monthly averages, we ignored individual missing days. If an entire month was missing, we filled in the values using the average of the previous and following months. After, we aligned the price and weather data to ensure the timestamps matched exactly. Finally, we split the dataset into a training set and a testing set using a 7:3 ratio.

## Time Series Models

In the early stage of modeling, we tested some time series models to capture the trend and seasonal patterns in cocoa prices. These models included ETS, ARIMA, SARIMA, and SARIMA-GARCH. To meet the stationarity requirement of time series models, we applied a log transformation and first-order differencing to the price data.

### ETS Models

To build the ETS model, we first plotted the time series of cocoa prices to examine long-term trends and seasonal patterns. For trend analysis, if the data showed a linear increase or decrease, we used an additive trend. If the trend appeared exponential—where higher price levels were linked to larger changes—we used a multiplicative trend.

Similarly, for seasonality, if seasonal fluctuations were roughly constant each year, we used an additive seasonal structure. If the size of seasonal changes grew or shrank with the price level, we used a multiplicative structure. However, certain components—such as the error term or unclear seasonal patterns—were hard to classify visually. For these cases, we used automated model selection to fit multiple ETS models and compared their corrected Akaike Information Criterion (AICc) values. We selected the model with the lowest AICc as the final ETS to balance model complexity and goodness of fit.

### ARIMA and SARIMA Models

For the ARIMA and SARIMA models, we set the differencing parameter d=0 because the data had already been transformed to achieve stationarity. We then plotted the autocorrelation function (ACF) and partial autocorrelation function (PACF) of the differenced series to examine the cutoff or tailing patterns of lags. This helped us make an initial judgment about the non-seasonal orders p and q, as well as the seasonal orders P and Q. For example, if the ACF drops sharply after a certain lag while the PACF tails off, it may suggest a strong moving average (MA) component and a weak autoregressive (AR) component, and vice versa.

When the ACF or PACF plots showed clear patterns, we used them to determine the non-seasonal order (p, d, q) and seasonal order (P, D, Q, s), then built the corresponding ARIMA or SARIMA models. However, in practice, the ACF and PACF of the price series often lacked clear cutoffs, and the decay of lags was vague, making it hard to identify optimal orders visually.

To address this, we used a grid search to fit models across a range of parameter combinations. We evaluated each model using the AICc and selected the model with the lowest AICc as the final ARIMA and SARIMA model, ensuring the best fit in the presence of structural complexity.

In order to improve model sensitivity to real-world factors, we included weather data as exogenous variables in the modeling process. These monthly variables, such as average temperature and precipitation, were aligned with the price data.

### SARIMA-GARCH Model

ARIMA and SARIMA models assume constant variance in their error terms. This assumption can lead to large prediction errors when the time series shows clear signs of conditional heteroskedasticity. To better capture potential volatility dynamics in the price series, we extended our analysis by building a SARIMA-GARCH model. This approach incorporates volatility modeling into the framework to improve stability and predictive performance. For the GARCH component, we used the commonly applied GARCH(1,1) specification.

## General Multiple Linear Regression Model

In addition to time series models, we also used a general multiple linear regression model based on weather variables to explore how weather affects cocoa prices from a different perspective. Since weather may influence prices with a time lag, we included lagged weather variables as predictors.

This approach helps capture delayed effects of weather on future prices and improves the model’s ability to respond to time dynamics. It also supports the assumption of independent error terms in linear regression by reducing the risk of autocorrelation.

To address the missing values created by lagged variables, we removed any records with incomplete data to ensure the regression was estimated using a full and valid sample. We also applied a log transformation to the cocoa price data. This reduced volatility, improved stability, and helped meet the normality and linearity assumptions of the model. As a result, the regression modeled the log of the expected price, log(E(Y|X)), rather than the raw expected price.

## Gradient Boosting Decision Tree Model

Building on the previous linear regression methods, we introduced a gradient boosting decision tree model to capture more complex, nonlinear relationships behind price fluctuations. This model combines multiple lagged values of historical prices as key time series features with weather variables—including precipitation, average temperature, maximum temperature, and minimum temperature—that are closely linked to price changes.

As with the linear model, we applied a log transformation to the price data to improve model stability and better align with the distribution of prediction errors.

During feature engineering, we created multiple lagged variables and aligned all weather features to ensure consistency. After preprocessing, we tuned the model by adjusting tree depth, learning rate, and the subsampling ratios for features and samples.

## Model Selection and Validation

### Model Selection Criteria

After fitting the models, we carried out evaluation and diagnostic analysis to assess both the fit on the training set and the generalization performance on the test set. For model selection, we used the AICc as the primary indicator. AICc balances model fit with complexity, helping to prevent overfitting. A lower AICc value indicates a better balance between complexity and fit, so we prioritized models with the lowest AICc scores.

For residual diagnostics, we examined whether the residuals met the white noise assumption. We plotted standardized residuals to check for randomness and to identify any signs of trend or structural bias. We also plotted the residual autocorrelation function (ACF) to see whether it cut off quickly after a small number of lags. Ideally, if the ACF shows no significant autocorrelation beyond lag 2, the model has likely captured most of the systematic information in the data.

We further conducted the Ljung-Box test. A p-value above the common threshold of 0.05 means we cannot reject the null hypothesis that the residuals are white noise, supporting the model’s validity.

For the linear regression model, we used a Q-Q plot to visually check if the residuals followed a normal distribution. We also calculated the variance inflation factor (VIF) for each predictor to assess multicollinearity. A VIF above 10 suggests a potential collinearity issue that may require variable removal or transformation. VIF values consistently below 5 indicate that the model has a stable variable structure.

### Prediction and Back-Transformation

After evaluating the models on the training set, we tested their predictive performance on the test set. It is important to note that different models applied different transformations to the cocoa price data during training. Therefore, we reversed these transformations before evaluation to ensure comparability of the results.

For the linear regression and XGBoost models, which were trained on the log-transformed prices, we applied an exponential transformation to the predictions to recover the actual price levels.

In contrast, the time series models—ETS, ARIMA, SARIMA, and SARIMA-GARCH—used both log transformation and first-order differencing to achieve stationarity. For these models, we first reconstructed the original log price series by cumulatively summing the differenced predictions. We then applied the exponential transformation to obtain the predicted actual prices.

### Evaluation Metrics and Final Model Selection

We evaluated model performance by using the optimal parameters estimated from the training set to predict cocoa prices in the test set. We then calculated several standardized prediction error metrics to assess accuracy: Mean Absolute Error (MAE), Root Mean Squared Error (RMSE), Mean Absolute Percentage Error (MAPE), and Sum of Squared Errors (SSE).

MAE measures the average absolute difference between predicted and actual values; lower values indicate better overall accuracy. RMSE gives more weight to larger errors, so a lower RMSE suggests the model performs well in avoiding large deviations. MAPE expresses prediction error as a percentage of the actual values, allowing for comparison across models and scales; lower values reflect smaller relative errors. SSE is the total of squared residuals and captures the model’s overall error; smaller values are preferred.

To avoid relying on a single metric, we compared all four indicators and paid particular attention to MAE and RMSE. Among the candidate models, we selected the one that showed consistent performance and relatively low errors across multiple metrics as the final prediction model. This model not only fit the training data well but also demonstrated strong and stable predictive accuracy on unseen data.

### 2025-04-04 liuchang/project/457.R

@@ -1,170 +1,0 @@
-# Load necessary libraries
-library(tidyverse)
-library(lubridate)
-library(forecast)
-library(tseries)
-
-# Load datasets
-cocoa_prices <- read.csv("00-data/Price.csv", stringsAsFactors = FALSE)
-ghana_weather <- read.csv("00-data/Climate.csv", stringsAsFactors = FALSE)
-
-# Convert date formats
-cocoa_prices$Date <- as.Date(cocoa_prices$Date, format='%d/%m/%Y')
-ghana_weather$DATE <- as.Date(ghana_weather$DATE)
-
-# Clean price data (remove commas and convert to numeric)
-cocoa_prices$Price <- as.numeric(gsub(",", "", cocoa_prices$ICCO.daily.price..US..tonne.))
-
-# Select and arrange data
-cocoa_prices <- cocoa_prices %>% select(Date, Price) %>% arrange(Date)
-
-# Aggregate weather data by date (mean values for daily observations)
-ghana_weather <- ghana_weather %>%
-  group_by(DATE) %>%
-  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))
-
-# Merge cocoa prices with weather data
-cocoa_data <- left_join(cocoa_prices, ghana_weather, by = c("Date" = "DATE"))
-cocoa_data <- na.omit(cocoa_data) # Remove missing values
-
-
-
-# 🟢 1️⃣ **Time Series Plot of Cocoa Prices**
-ggplot(cocoa_data, aes(x = Date, y = Price)) +
-  geom_line(color = "blue", size = 1) +
-  labs(title = "Cocoa Price Over Time",
-       x = "Date", 
-       y = "Cocoa Price (USD/tonne)") +
-  theme_minimal()
-library(corrplot)
-library(gridExtra)
-# 🟢 2️⃣ **Histogram & Density Plot of Cocoa Prices**
-p1 <- ggplot(cocoa_data, aes(x = Price)) +
-  geom_histogram(fill = "blue", bins = 30, alpha = 0.7) +
-  labs(title = "Histogram of Cocoa Prices",
-       x = "Cocoa Price (USD/tonne)", 
-       y = "Frequency") +
-  theme_minimal()
-
-p2 <- ggplot(cocoa_data, aes(x = Price)) +
-  geom_density(fill = "blue", alpha = 0.5) +
-  labs(title = "Density Plot of Cocoa Prices",
-       x = "Cocoa Price (USD/tonne)", 
[…]

### 2025-04-04 liuchang/流畅的最终版本.R

@@ -1,266 +1,0 @@
-### Load required libraries
-library(tidyverse)
-library(lubridate)
-library(forecast)
-library(tseries)
-library(ggplot2)
-library(xgboost)
-library(caret)
-library(slider)
-library(rugarch)
-
-### Load and Aggregate Cocoa Price Data to Monthly
-cocoa_prices <- read.csv("Daily Prices_ICCO.csv", stringsAsFactors = FALSE)
-cocoa_prices$Date <- as.Date(cocoa_prices$Date, format='%d/%m/%Y')
-cocoa_prices$Price <- as.numeric(gsub(",", "", cocoa_prices$ICCO.daily.price..US..tonne.))
-cocoa_prices <- cocoa_prices %>%
-  mutate(YearMonth = floor_date(Date, "month")) %>%
-  group_by(YearMonth) %>%
-  summarise(Price = mean(Price, na.rm = TRUE)) %>%
-  ungroup()
-
-### Load and Aggregate Ghana Weather Data to Monthly
-ghana_weather <- read.csv("Ghana_data.csv", stringsAsFactors = FALSE)
-ghana_weather$DATE <- as.Date(ghana_weather$DATE)
-ghana_weather <- ghana_weather %>%
-  mutate(YearMonth = floor_date(DATE, "month")) %>%
-  group_by(YearMonth) %>%
-  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))
-
-### Merge and Clean Monthly Data
-cocoa_data <- left_join(cocoa_prices, ghana_weather, by = "YearMonth") %>%
-  mutate(log_price = log(Price),
-         diff_log_price = c(NA, diff(log_price))) %>%
-  drop_na()
-
-### Plot Monthly Time Series
-ggplot(cocoa_data, aes(x = YearMonth)) +
-  geom_line(aes(y = Price), color = "steelblue") +
-  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
-  theme_minimal()
-
-### Split Data into Training and Testing Sets
-train_size <- floor(0.8 * nrow(cocoa_data))
-train_data <- cocoa_data[1:train_size, ]
-test_data <- cocoa_data[(train_size + 1):nrow(cocoa_data), ]
-
-### Fit ETS and ARIMA Models
-ets_model_1 <- ets(train_data$diff_log_price)
-ets_model_2 <- ets(train_data$diff_log_price, model = "ZZZ")
[…]

### 2025-04-04 Final code.Rmd
Add files via upload

---
title: "STA457 Final Project"
author: "Wenyi Li"
date: "2025-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
# load required libraries
library(tidyverse)
library(lubridate)
library(forecast)
library(tseries)
library(ggplot2)
library(xgboost)
library(caret)
library(slider)
library(rugarch)
library(car)
```

# Data Preprocessing
```{r}
# load and preprocess price data
cocoa_prices <- read.csv("Daily Prices_ICCO.csv", stringsAsFactors = FALSE)
cocoa_prices$Date <- as.Date(cocoa_prices$Date, format='%d/%m/%Y')
cocoa_prices$Price <- as.numeric(gsub(",", "", cocoa_prices$ICCO.daily.price..US..tonne.))
cocoa_prices <- cocoa_prices %>%
  mutate(YearMonth = floor_date(Date, "month")) %>%
  group_by(YearMonth) %>%
  summarise(Price = mean(Price, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
# load and preprocess Ghana weather data
ghana_weather <- read.csv("Ghana_data.csv", stringsAsFactors = FALSE)
ghana_weather$DATE <- as.Date(ghana_weather$DATE)
ghana_weather <- ghana_weather %>%
  mutate(YearMonth = floor_date(DATE, "month")) %>%
  group_by(YearMonth) %>%
  summarise(across(c(PRCP, TAVG, TMAX, TMIN), mean, na.rm = TRUE))
```

```{r}
# Merge and Clean Monthly Data(log + diff)
cocoa_data <- left_join(cocoa_prices, ghana_weather, by = "YearMonth") %>%
  mutate(log_price = log(Price),
         diff_log_price = c(NA, diff(log_price))) %>%
  drop_na()

# Plot Monthly Time Series
ggplot(cocoa_data, aes(x = YearMonth)) +
  geom_line(aes(y = Price), color = "steelblue") +
  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal()
```

```{r}
# Split Data into Training and Testing Sets(7:3 ratio)
train_size <- floor(0.7 * nrow(cocoa_data))
train_data <- cocoa_data[1:train_size, ]
test_data <- cocoa_data[(train_size + 1):nrow(cocoa_data), ]
```

# Modeling
```{r}
# since not station, already transformed in the pre-processed part(log + diff)
# built ets models
ets_model_1 <- ets(train_data$diff_log_price, model = "ZZZ")
ets_model_2 <- ets(train_data$diff_log_price)
plot(ts(fitted(ets_model_2)), col = "red", 
     main = "Actual vs Predicted for diff-log price ets") + 
  lines(ts(train_data$diff_log_price), col = "green")
# ets_model_2(ets_model_1) was selected as a candidate
```

```{r}
# built arima, sarima models
# verify stationarity(1st diff + log)
ggplot(train_data, aes(x = YearMonth)) +
  geom_line(aes(y = diff_log_price), color = "steelblue") +
  labs(title = "Monthly Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal()
acf(train_data$diff_log_price, main = "ACF of differencing log price")
pacf(train_data$diff_log_price, main = "PACF of differencing log price")
external_regressors <- data.matrix(train_data[, c("PRCP", "TAVG", "TMAX", "TMIN")])
arima_model1 <- arima(train_data$diff_log_price, order = c(2,0,2), xreg = external_regressors)
arima_model2 <- arima(train_data$diff_log_price, order = c(2,0,5), xreg = external_regressors)
sarimax_model <- auto.arima(train_data$diff_log_price, xreg = as.matrix(external_regressors), seasonal = TRUE)
summary(arima_model1)
summary(arima_model2)
# arima_model1, sarimax_model was selected as a candidate, seasonality not detected(sarimaxmodel same as arimamodel1)
plot(ts(fitted(arima_model1)), col = "red", main = "Actual vs Predicted for diff-log price arima") + lines(ts(train_data$diff_log_price), col = "green")
plot(ts(fitted(sarimax_model)), col = "red", main = "Actual vs Predicted for diff-log price sarimax") + lines(ts(train_data$diff_log_price), col = "green")
```

```{r}
# residual diagnostic for arima, sarima models
tsdiag(arima_model1, gof.lag = 20)
tsdiag(arima_model2, gof.lag = 20)
tsdiag(sarimax_model, gof.lag = 20)
```

```{r}
# forecasting in diff-log base
test_xreg <- data.matrix(test_data[, c("PRCP", "TAVG", "TMAX", "TMIN")])
test_xreg <- as.matrix(test_xreg)
ets_forecast_2 <- forecast(ets_model_2, h = nrow(test_data))
sarimax_forecast <- forecast(sarimax_model, xreg = test_xreg, h = nrow(test_data))
h <- nrow(test_data)
pred1 <- predict(arima_model1, n.ahead = h, newxreg = test_xreg)
```

```{r}
# back-transform forecasted values
reconstruct_log_prices <- function(last_log_price, diffs) {cumsum(c(last_log_price, diffs))[-1]}

last_log_price <- tail(train_data$log_price, 1)
n <- nrow(test_data)
forecast_dates <- test_data$YearMonth

ets2_log_forecast <- reconstruct_log_prices(last_log_price, ets_forecast_2$mean)
sarimax_log_forecast <- reconstruct_log_prices(last_log_price, sarimax_forecast$mean)
arima_log_forecast <- reconstruct_log_prices(last_log_price, pred1$pred)

ets2_price_forecast <- exp(ets2_log_forecast)
ets2_price_forecast
sarimax_price_forecast <- exp(sarimax_log_forecast)
sarimax_price_forecast
arima_price_forecast <- exp(arima_log_forecast)
arima_price_forecast

forecast_df <- bind_rows(
  tibble(Date = forecast_dates, Forecast = ets2_price_forecast, Model = "ETS Model 2"),
  tibble(Date = forecast_dates, Forecast = sarimax_price_forecast, Model = "SARIMAX"),
  tibble(Date = forecast_dates, Forecast = arima_price_forecast, Model = "ARIMA")
) %>% drop_na()

forecast_df
#forecast_df is a knitted back-transformed forecasted values for three models

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.3) +
  geom_line(data = forecast_df, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.3) +
  labs(title = "Monthly Predicted of Three Models vs Actual Cocoa Prices", y = "Price", x = "Date") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c(
    "ETS Model 2" = "green",
    "SARIMAX" = "blue",
    "ARIMA" = "red"
  )) + 
  scale_linetype_manual(values = c(
    "ETS Model 2" = "solid",
    "SARIMAX" = "dotdash",
    "ARIMA" = "twodash"
  ))

# acc comparison
actual_prices <- exp(test_data$log_price)
ets2_acc <- accuracy(ets2_price_forecast, actual_prices)
sarimax_acc <- accuracy(sarimax_price_forecast, actual_prices)
arima_acc <- accuracy(arima_price_forecast, actual_prices)

print("ETS Model Performance:")
print(ets2_acc)
print("ARIMAX Model Performance:")
print(arima_acc)
print("SARIMAX Model Performance:")
print(sarimax_acc)
```

```{r}
# built linear regression model
# create Lag Features
generate_lags <- function(data, lags = 1:6) {
  for (lag in lags) {
    data[[paste0("lag_", lag)]] <- dplyr::lag(data$log_price, lag)
  }
  return(data)
}
cocoa_data_lagged <- generate_lags(cocoa_data) %>% drop_na()
lm_data <- cocoa_data_lagged %>%
  select(YearMonth, log_price, starts_with("lag_"), PRCP, TAVG, TMAX, TMIN)
train_size <- floor(0.7 * nrow(lm_data))
train_lm <- lm_data[1:train_size, ]
test_lm <- lm_data[(train_size + 1):nrow(lm_data), ]
lm_model <- lm(log_price ~ ., data = train_lm %>% select(-YearMonth))
plot(ts(fitted(lm_model)), col = "red", 
     main = "Actual vs Predicted for log price regression") + 
  lines(ts(train_data$log_price), col = "green")

lm_pred_log <- predict(lm_model, newdata = test_lm)
lm_pred_price <- exp(lm_pred_log)
lm_results <- tibble(
  Date = test_lm$YearMonth,
  Actual = exp(test_lm$log_price),
  Predicted = lm_pred_price
)
### Plot Regression Results
ggplot(lm_results, aes(x = Date)) +
  geom_line(aes(y = Actual), color = "red") +
  geom_line(aes(y = Predicted), color = "blue") +
  labs(title = "Linear Regression Predicted vs Actual Prices (Monthly)", y = "Price", x = "Date") +
  theme_minimal()

lm_accuracy <- accuracy(lm_pred_price, exp(test_lm$log_price))
print("Linear Regression Model Performance:")
print(lm_accuracy)
```

```{r}
acf(residuals(lm_model), main = "ACF of Residuals regression")
library(lmtest)
dwtest(lm_model) #p-value > 0.05 good -> no autocorrelation
qqnorm(residuals(lm_model), main = "QQ-Plot of regression residuals")
qqline(residuals(lm_model), col = "red", lwd = 2)
vif(lm_model)
Box.test(residuals(lm_model), lag = 20, type = "Ljung-Box")
```

```{r}
# Calculate log returns
log_re <- diff(log(cocoa_data$Price))
log_re <- na.omit(log_re)
train_size <- floor(0.7 * length(log_re))
train_re <- log_re[1:train_size]
test_re <- log_re[(train_size + 1):length(log_re)]
test_dates <- cocoa_data$YearMonth[(train_size + 2):(length(log_re) + 1)]
# define garch model - using our best arma model p,q with 2,2 and widely used garchorder 1,1
garch_spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(2, 2), include.mean = TRUE),
  distribution.model = "norm"
)
garch_fit <- ugarchfit(spec = garch_spec, data = train_re)
garch_forecast <- ugarchforecast(garch_fit, n.ahead = length(test_re))
predicted_re <- as.numeric(fitted(garch_forecast))
plot(ts(fitted(garch_fit)), col = "red", 
     main = "Actual vs Predicted for diff-log price garch") +
  lines(ts(train_data$diff_log_price), col = "green")

# back-transform
last_train_price <- cocoa_data$Price[train_size + 1]
forecast_prices <- last_train_price * exp(cumsum(predicted_re))

garch_df <- tibble(
  Date = test_dates,
  Price = forecast_prices
)
# plot the model
ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black") +
  geom_line(data = garch_df, aes(x = Date, y = Price), color = "pink") +
  labs(title = "GARCH Predicted vs Actual Prices (Monthly)", y = "Price", x = "Date") +
  theme_minimal()

garch_accuracy <- accuracy(forecast_prices, exp(test_data$log_price))

print("GARCH Model Performance:")
print(garch_accuracy)
```
```{r}
residuals_std <- residuals(garch_fit, standardize = TRUE)
acf(residuals_std, main = "ACF of Residuals garch")
Box.test(residuals_std, lag = 20, type = "Ljung-Box")

qqnorm(residuals_std, main = "QQ-Plot of ARMA-GARCH")
qqline(residuals_std, col = "red", lwd = 2)
```

```{r}
# built xgboost model
# generate lags
generate_lags <- function(data, lags = 1:6) {
  for (lag in lags) {
    data[[paste0("lag_", lag)]] <- dplyr::lag(data$log_price, lag)
  }
  return(data)
}

cocoa_data_lagged <- cocoa_data %>%
  generate_lags() %>%
  drop_na()

initial_size <- floor(0.7 * nrow(cocoa_data_lagged))
forecast_horizon <- nrow(cocoa_data_lagged) - initial_size

walk_results <- map_dfr(1:forecast_horizon, function(i) {
  train_set <- cocoa_data_lagged[1:(initial_size + i - 1), ]
  test_point <- cocoa_data_lagged[(initial_size + i), ]

  x_train <- train_set %>% select(starts_with("lag_"), PRCP, TAVG, TMAX, TMIN)
  y_train <- train_set$log_price
  x_test <- test_point %>% select(starts_with("lag_"), PRCP, TAVG, TMAX, TMIN)

  dtrain <- xgb.DMatrix(as.matrix(x_train), label = y_train)
  dtest <- xgb.DMatrix(as.matrix(x_test))
# fit model
  model <- xgboost(
    data = dtrain,
    nrounds = 600,
    objective = "reg:squarederror",
    verbose = 0,
    eta = 0.05
  )

  pred_log <- predict(model, dtest)
  tibble(
    Date = test_point$YearMonth,
    Actual = exp(test_point$log_price),
    Predicted = exp(pred_log)
  )
})

ggplot(walk_results, aes(x = Date)) +
  geom_line(aes(y = Actual), color = "black", linewidth = 1) +
  geom_line(aes(y = Predicted), color = "purple", linewidth = 1) +
  labs(title = "XGBoost Walk-Forward Forecast vs Actual (Monthly)",
       y = "Price", x = "Date") +
  theme_minimal()

xgb_accuracy <- accuracy(walk_results$Predicted, walk_results$Actual)
print("XGBoost Walk-Forward Accuracy Metrics:")
print(xgb_accuracy)
```

```{r}
# residual analysis of xgboost model
walk_results <- walk_results %>%
  filter(!is.na(Actual), !is.na(Predicted)) %>%
  mutate(Residual = Actual - Predicted)
acf(walk_results$Residual, main = "ACF of Residuals xgboost")
Box.test(walk_results$Residual, lag = 20, type = "Ljung-Box")

x_train_full <- cocoa_data_lagged %>% select(starts_with("lag_"), PRCP, TAVG, TMAX, TMIN)
y_train_full <- cocoa_data_lagged$log_price
dtrain_full <- xgb.DMatrix(as.matrix(x_train_full), label = y_train_full)

final_model <- xgboost(
  data = dtrain_full,
  nrounds = 600,
  objective = "reg:squarederror",
  eta = 0.05,
  verbose = 0
)
```

```{r}
h_future <- 120
ets_forecast_future <- forecast(ets_model_2, h = h_future)

last_log_price <- tail(train_data$log_price, 1)
ets_log_future <- reconstruct_log_prices(last_log_price, ets_forecast_future$mean)

# backtransform
ets_price_future <- exp(ets_log_future)

future_dates <- seq.Date(from = max(cocoa_data$YearMonth) + months(1),
                         by = "month", length.out = h_future)

ets_forecast_df_10yr <- tibble(
  Date = future_dates,
  Forecast = ets_price_future,
  Model = "ETS Model 2"
)

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = ets_forecast_df_10yr, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices by ETS Model",
       x = "Date", y = "Price") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
# calculate historical averages from training data
avg_PRCP  <- mean(train_data$PRCP, na.rm = TRUE)
avg_TAVG  <- mean(train_data$TAVG, na.rm = TRUE)
avg_TMAX  <- mean(train_data$TMAX, na.rm = TRUE)
avg_TMIN  <- mean(train_data$TMIN, na.rm = TRUE)

future_xreg <- matrix(rep(c(avg_PRCP, avg_TAVG, avg_TMAX, avg_TMIN),
                          each = h_future), nrow = h_future)

arima_future <- predict(arima_model1, n.ahead = h_future, newxreg = future_xreg)

# back-transform
arima_log_future <- reconstruct_log_prices(last_log_price, arima_future$pred)
arima_price_future <- exp(arima_log_future)

arima_forecast_df_10yr <- tibble(
  Date = future_dates,
  Forecast = arima_price_future,
  Model = "ARIMA"
)

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = arima_forecast_df_10yr, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices by ARIMA Model",
       x = "Date", y = "Price") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
future_lm_preds <- numeric(h_future)

current_lags <- tail(lm_data$log_price, 6)

for(i in 1:h_future) {
  newdata <- as.data.frame(t(c(current_lags, avg_PRCP, avg_TAVG, avg_TMAX, avg_TMIN)))
  colnames(newdata) <- c(paste0("lag_", 1:6), "PRCP", "TAVG", "TMAX", "TMIN")
  pred_log <- predict(lm_model, newdata = newdata)
  future_lm_preds[i] <- pred_log

  current_lags <- c(pred_log, current_lags[1:5])
}

lm_price_future <- exp(future_lm_preds)

lm_forecast_df_10yr <- tibble(
  Date = future_dates,
  Forecast = lm_price_future,
  Model = "Linear Regression"
)

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = lm_forecast_df_10yr, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices by Linear Regression Model",
       x = "Date", y = "Price") +
  theme_minimal() +
  theme(legend.position = "bottom")

write.csv(lm_forecast_df_10yr, "output_file.csv", row.names = FALSE)
```

```{r}
garch_forecast_10yr <- ugarchforecast(garch_fit, n.ahead = 120)
predicted_10yr_returns <- as.numeric(fitted(garch_forecast_10yr))

last_price <- tail(cocoa_data$Price, 1)
future_garch_prices <- last_price * exp(cumsum(predicted_10yr_returns))

garch_forecast_df_10yr <- tibble(
  Date = future_dates,
  Forecast = future_garch_prices,
  Model = "GARCH"
)

ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = garch_forecast_df_10yr, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices by ARMA-GARCH Model",
       x = "Date", y = "Price") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
h_future <- 120
future_xgb_preds <- numeric(h_future)

# Get last 6 log prices (for lag_1 to lag_6)
current_lags <- as.numeric(tail(cocoa_data_lagged$log_price, 6))

# Average weather values (for fixed future input)
avg_PRCP <- mean(cocoa_data_lagged$PRCP, na.rm = TRUE)
avg_TAVG <- mean(cocoa_data_lagged$TAVG, na.rm = TRUE)
avg_TMAX <- mean(cocoa_data_lagged$TMAX, na.rm = TRUE)
avg_TMIN <- mean(cocoa_data_lagged$TMIN, na.rm = TRUE)

# Forecast loop
for (i in 1:h_future) {
  # Construct input as data frame
  newdata <- as.data.frame(t(c(current_lags, avg_PRCP, avg_TAVG, avg_TMAX, avg_TMIN)))
  colnames(newdata) <- c(paste0("lag_", 1:6), "PRCP", "TAVG", "TMAX", "TMIN")

  # Convert to matrix, then to DMatrix
  newdata_matrix <- as.matrix(newdata)
  dnew <- xgb.DMatrix(newdata_matrix)

  # Predict log price
  pred_log <- predict(final_model, dnew)

  # Save prediction
  future_xgb_preds[i] <- pred_log

  # Update lags
  current_lags <- c(pred_log, current_lags[1:5])
}

# Convert log-price to price
xgb_price_future <- exp(future_xgb_preds)

# Generate future dates
library(lubridate)
last_date <- max(cocoa_data_lagged$YearMonth)
future_dates <- seq.Date(from = last_date %m+% months(1), by = "month", length.out = h_future)

# Final forecast data frame
xgb_forecast_df_10yr <- tibble(
  Date = future_dates,
  Forecast = xgb_price_future,
  Model = "XGBoost"
)

# Plot
ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = xgb_forecast_df_10yr, aes(x = Date, y = Forecast, color = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices (XGBoost)",
       x = "Date", y = "Price") +
  theme_minimal()
```
```{r}
# Combine all forecast data frames
all_forecasts_10yr <- bind_rows(
  ets_forecast_df_10yr,
  arima_forecast_df_10yr,
  lm_forecast_df_10yr,
  garch_forecast_df_10yr,
  xgb_forecast_df_10yr
)

# Plot the forecasts alongside historical prices
ggplot() +
  geom_line(data = cocoa_data, aes(x = YearMonth, y = Price), color = "black", linewidth = 1.2) +
  geom_line(data = all_forecasts_10yr, aes(x = Date, y = Forecast, color = Model, linetype = Model), linewidth = 1.2) +
  labs(title = "10-Year Forecast of Cocoa Prices by Model",
       x = "Date", y = "Price") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
